Cycle Route Uptake and Scenario Estimation (CRUSE): an open and nationally scalable approach for strategic cycle network planning tool development

### 2024-03-23 paper.qmd
some edits

# # Uncomment the following line to produce the cover page
author:
  - name: Robin Lovelace
    email: r.lovelace@leeds.ac.uk
    corresponding: true
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Joey Talbot
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Eugeni Vidal-Tortosa
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Hussein Mahfouz
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Elaine Brick
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Peter Wright
    affiliations:
      - name: AECOM
        address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY, UK
  - name: "Gary O'Tool"
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Dan Brennan
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Suzanne Meade
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  This paper describes an approach to the development of strategic cycle network planning tools. Based on our experience developing and deploying the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we outline the methods used to generate the evidence presented in the tool, including disaggregation of origin-destination data, additional trip purposes, routing, scenarios of cycling potential, and the user interface. The CRUSE Tool provides estimates of cycling currently and under four scenarios of cycling potential: Near Market, Climate Action Plan, Go Dutch, and Ebike. The tool was commissioned by Transport Infrastructure Ireland and is publicly available at <https://cruse.bike/>, enabling planners, engineers, and other stakeholders to make more evidence-based decisions. CRUSE goes beyond previous work by: modelling networks at high levels of geographic resolution; simulating multiple trip purposes (social, shopping and personal utility trips plus recreational and cycle touring trips), in addition to using official data on trips to work and education locations; provision of estimates of 'quietness' (a proxy for cyclability) at the route segment level; and its provision of three network types: 'Fastest', 'Balanced' and 'Quietest' to help plan both arterial and residential cycle networks. Workshops and unstructure interview is stakeholders show that the tool is already being used in practice, to inform urban, inter-urban and rural network development. The approach and methods for strategic cycle network development tools outlined in this paper can be applied and built-on in other places to support the generation of evidence needed to support ambitious cycling policies and interventions internationally.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
date: last-modified
---

@@

```{r}
#| include: false
#| eval: false
remotes::install_dev("tmap")
library(tmap)
# Save title page (1st page of paper.pdf):
system("pdftk _site/paper.pdf cat 1 output _site/title_page.pdf")
# Combine cv.pdf with visions.pdf:
system("pdftk _site/cv.pdf _site/visions.pdf cat output _site/cv_and_visions.pdf")
# latexdiffr::latexdiff("paper-v1.qmd", "paper.qmd")
system("latexdiff paper-v1.tex paper.tex > paper-diff.tex")
system("pdflatex paper-diff.tex")
browseURL("paper-diff.pdf")
```

@@

Growing evidence of the negative impacts of car-dependent transport systems has led governments in many countries to set targets and taken actions.
In the context of climate, road safety and physicial inactivity crises, policies to improve transport systems can be classified according to the 'Avoid-Shift-Improve' (ASI) framework [@jaramillo2022].
The ASI framework highlights the importance of demand reduction (*avoid*ing unnecessary trips), in addition to mode *shift* to sustainable modes and *improvement* of existing energy converters, in that order.

@@

Another motivation for cycle network planning at the European level is the Road Infrastructure Safety Management (RISM) directive (2008/96/EC), which requires member states to implement a road safety management system (RSMS) for all public roads.
Specifically, "Member States shall ensure that the ranking of high accident concentration sections and the network safety ranking are carried out" [@directiv2008].
Given that 'safety' in this context is usefully quantified as the number of people killed and seriously injured (KSI) per distance traveled, the directive requires estimation of distance traveled by mode, down to the road link level.
For active modes, about which there is a paucity of data compared with motorized modes, this is a major challenge.
Better data to inform road safety policies and interventions is a motivation for better estimates of 'baseline' levels of physical activity at high levels of geographic resolution [@tait2023].

@@

Stakeholder consultation by TII emphasized the need for a strong, national, systematic but locally-specific evidence base on cycle networks in both urban and rural Ireland.
The evidence base had to scale nationally to support strategic alignment with national, regional and local policies, but also be useful for local network planning.
TII also requires

Illustration of the density of route networks generated by the CRUSE Tool for Dublin City and surroundings (top) and rural County Mayo on the West coast of Ireland (botton).
Source: CRUSE Tool, available at <https://cruse.bike/>.
:::

This paper aims to describe the design, features and potential use of the CRUSE Tool.
In @sec-methods, we outline the methods used to generate the evidence presented in the CRUSE Tool.
In @sec-results, we present the results of the CRUSE Tool, including estimates of current cycling levels and future cycling potential at the national, regional and local levels.
In @sec-discussion, we discuss the implications of the results for policy and practice.
In @sec-conclusions, we conclude the paper.

@@

@@

The aim of the paper was to describe the design, features and potential use of the CRUSE Tool.
As outlined in the previous section, the CRUSE Tool provides evidence on current and potential future cycling levels across Ireland down to the street level, with potential assigned to 'fastest', 'balanced' and 'quietest' route networks.
This provision of multiple scenarios of behavior change *and* multiple scenarios of investment, for example in cycling infrastructure next to major roads vs quiet residential streets, is a key feature of the CRUSE Tool.

@@

>>>>

# # Uncomment the following line to produce the cover page
author:
  - name: Robin Lovelace
    email: r.lovelace@leeds.ac.uk
    corresponding: true
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Joey Talbot
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Eugeni Vidal-Tortosa
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Hussein Mahfouz
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Elaine Brick
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Peter Wright
    affiliations:
      - name: AECOM
        address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY, UK
  - name: "Gary O'Toole"
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Dan Brennan
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Suzanne Meade
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  This paper describes an approach to the development of strategic cycle network planning tools. Based on our experience developing and deploying the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we outline the methods used to generate the evidence presented in the tool, including disaggregation of origin-destination data, additional trip purposes, routing, scenarios of cycling potential, and the user interface. The CRUSE Tool provides estimates of cycling currently and under four scenarios of cycling potential: Near Market, Climate Action Plan, Go Dutch, and Ebike. The tool was commissioned by Transport Infrastructure Ireland and is publicly available at <https://cruse.bike/>, enabling planners, engineers, and other stakeholders to make more evidence-based decisions. CRUSE goes beyond previous work by: modelling networks at high levels of geographic resolution; simulating multiple trip purposes (social, shopping and personal utility trips plus recreational and cycle touring trips), in addition to using official data on trips to work and education locations; provision of estimates of 'quietness' (a proxy for cyclability) at the route segment level; and its provision of three network types: 'Fastest', 'Balanced' and 'Quietest' to help plan both arterial and residential cycle networks. Workshops and unstructured interviews of stakeholders show that the tool is already being used in practice, to inform urban, inter-urban and rural network development. The approach and methods for strategic cycle network development tools outlined in this paper can be applied and built-on in other places to support the generation of evidence needed to support ambitious cycling policies and interventions internationally.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
date: last-modified
---

@@

```{r}
#| include: false
#| eval: false
remotes::install_dev("tmap")
library(tmap)
# Save title page (1st page of paper.pdf):
system("pdftk _site/paper.pdf cat 1 output _site/title_page.pdf")
# Combine cv.pdf with visions.pdf:
system("pdftk _site/cv.pdf _site/visions.pdf cat output _site/cv_and_visions.pdf")
# latexdiffr::latexdiff("paper-v1.qmd", "paper.qmd")
system("latexdiff paper-v1.tex paper.tex > tracked-changes.tex")
system("pdflatex tracked-changes.tex")
browseURL("tracked-changes.pdf")
```

@@

Growing evidence of the negative impacts of car-dependent transport systems has led governments in many countries to set targets and take actions.
In the context of climate, road safety and physicial inactivity crises, policies to improve transport systems can be classified according to the 'Avoid-Shift-Improve' (ASI) framework [@jaramillo2022].
The ASI framework highlights the importance of demand reduction (*avoid*ing unnecessary trips), in addition to mode *shift* to sustainable modes and *improvement* of existing energy converters, in that order.

@@

Another motivation for cycle network planning at the European level is the Road Infrastructure Safety Management (RISM) directive (2008/96/EC), which requires member states to implement a road safety management system (RSMS) for all public roads.
Specifically, "Member States shall ensure that the ranking of high accident concentration sections and the network safety ranking are carried out" [@directiv2008].
Given that 'safety' in this context is usefully quantified as the number of people killed or seriously injured (KSI) per distance traveled, the directive requires estimation of distance traveled by mode, down to the road link level.
For active modes, about which there is a paucity of data compared with motorized modes, this is a major challenge.
Better data to inform road safety policies and interventions is a motivation for better estimates of 'baseline' levels of physical activity at high levels of geographic resolution [@tait2023].

@@

Stakeholder consultation by TII emphasized the need for a strong, national, systematic but locally-specific evidence base on cycle networks in both urban and rural Ireland.
The evidence base had to scale nationally to support strategic alignment with national, regional and local policies, but also be useful for local network planning.

::: {#fig-landing}
![](images/paste-10.png)

Illustration of the landing page of the CRUSE Tool for Ireland.
:::

This describes describe the design, features and potential use of an approach to strategic cycle network planning that is open, scalable, and evidence-based.
We label the approach Cycle Route Uptake and Scenario Estimation (CRUSE) and present a case study of its application in Ireland, the results of which are publicly available at [cruse.bike](https://cruse.bike/) in the CRUSE Tool for Ireland, the landing page of which is illustrated in @fig-landing.
In @sec-methods, we outline the methods used to generate the evidence presented in the CRUSE Tool.
In @sec-results, we present the results of the CRUSE Tool, including estimates of current cycling levels and future cycling potential at the national, regional and local levels.
In @sec-discussion, we discuss limitations and possible future improvements to the approach, and the implications of the results for Ireland before making concluding remarks in @sec-conclusions.

@@

Illustration of the density of route networks generated by the CRUSE Tool for Dublin City and surroundings (top) and rural County Mayo on the West coast of Ireland (botton).
Source: CRUSE Tool, publicly available at [cruse.bike](https://cruse.bike/).
:::

@@

The aim of the paper was to describe the design, features and potential use of the 'CRUSE' approach and resulting tool.
As outlined in the previous section, the CRUSE Tool provides evidence on current and potential future cycling levels across Ireland down to the street level, with potential assigned to 'fastest', 'balanced' and 'quietest' route networks.
This provision of multiple scenarios of behavior change *and* multiple scenarios of investment, for example in cycling infrastructure next to major roads vs quiet residential streets, is a key feature of the CRUSE Tool.

## Uses of the tool in practice

@@

## Comparison with other tools and contribution to the field
